---
title: "Week2 Async"
author: "Jenn Dandrea"
date: "10/3/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
```{r}
#read in raw data
titantic<-read.csv("/Users/jenndand/Desktop/Syracuse/707/707R/TitanticRaw.csv")
str(titantic)

#2.5, converting passenger ID to factor
titantic$PassengerId<-factor(titantic$PassengerId)
str(titantic$PassengerId)
```
2.6
The ages had the most missing values.I'm going to omit the age where they are missing. because I dont want to throw off the data if I want to caluate average age of people, but maybe the sex/class is critical, so i dont want to remove the whole rows of data.

```{r}
#2.6 how many missing items
length(which(is.na(titantic)))
#87 missing values
nrow(titantic[!complete.cases(titantic),])
#87 rows are not complete
nrow(titantic[complete.cases(titantic),])
#331 rows are complete

length(which(is.na(titantic$Age)))
#86 rows are missing an age values, and 87 (above) had not complete rows. 
table(is.na(titantic)) #provides how many entries do not have NA in them and how many do
sapply(titantic, function(x) sum(is.na(x))) #provides a count of n/a by each column
titantic[!complete.cases(titantic),] #Shows all rows with n/a in them

titantic$Fare[is.na(titantic$Fare)] <- mean(titantic$Fare, na.rm = TRUE) #replacing the Fare n/a with an average of the fare prices
titantic <- na.omit(titantic) #omiting the na in the data


```

2.7
```{r}
#mean = 2.2
mean(titantic$Pclass) 
# median = 3
median(titantic$Pclass)
#mode 3
table(titantic$Pclass)[which.max(table(titantic$Pclass))] # mode

#creating a table of the frequence, 1=107, 2=93, 3=218. Most people were in the 3 class
freq<-table(titantic$Pclass) # frequency distribution
freq

table(titantic$Pclass)[which.max(table(titantic$Pclass))] # mode
#Variance = .70
var(titantic$Pclass) # variance
#standard deviation =.84
sd(titantic$Pclass) # standard deviation
#max = 3
max(titantic$Pclass)
#min =1
min(titantic$Pclass)

```
2.8.1
```{r}
# Histogram of the PClass
#Creating smaller subset with just the pclass=3
Pclass3<- titantic[titantic$Pclass == "3",]
boxplot(Pclass3$Fare,  main="Pclass 3 v Fare", xlab="Pclass 3", ylab="Fare")
#There are 9 outliers in this data


hist(Pclass3$Fare) #, main="Pclass v Fare", xlab="Pclass") 

# Boxplot for all pClasses
boxplot(Fare~Pclass,data=titantic, main="Pclass v Fare", xlab="Pclass", ylab="Fare") 


#Crosstab of the embark locations and if the passengered survived (1 for=surivied, 0=died)
EmbarSurvived<-table(titantic$Embarked, titantic$Survived)
EmbarSurvived
#corrolation between the S location and non-surviving is high. Do an additional cross tab of the Pclass and the embarked, 142 people from S were Pclass=3. So they were in the lower class, and therefor more probable to die.

PClassSurvived1<-table(titantic$Pclass,  titantic$Embarked)
PClassSurvived1
```

2.9 Aggregation the average fare for men and for women in the Titanic data set.
```{r}

attach(titantic)

aggregate(Fare~Sex, data=titantic, FUN=mean)

```
2.10.1 Exercise: Transformation Choose a numeric variable in the Titanic data set and transform it using discretization, log, z-score, and min-max transformation.
```{r}
#Selecting Age since it is a numberic value in the data
#calculating the Z score
scale(titantic$Age, center = TRUE, scale = TRUE)

#Discretization
age <- cut(titantic$Age, breaks =c(0,10,20,30,40,50,60,Inf),labels=c("child","teens","twenties","thirties","fourties","fifties","old"))
table(age)
#most people where in their 20s

#Log transformation
plot(titantic$Age, log(titantic$Age))

#Min-max transformation 
Min_max <-(titantic$Age-min(titantic$Age,na.rm=TRUE))/(max(titantic$Age,na.rm=TRUE)-min(titantic$Age,na.rm=TRUE))

plot(Min_max, titantic$Age)

```
2.11 Produce a random sample of 100 examples out of the Titanic data. Produce a systematic sample of 100 examples out of the Titanic data. Show your R code.
```{r}
sample <- titantic[sample(1:nrow(titantic), 100, replace=FALSE), ]
# pulling 100 samples with out replacement. 

#View(sample) this will show all the samples, which I dont want to do right now
table(sample$Survived) #table of the sample for survived and did not 
table(titantic$Survived) #table of the raw data of survived and did not


#Systematic sampling

syssample<-titantic[seq(1, nrow(titantic),22),] # sample every 22 lines
syssample

```
